---
title: "RVC for Infographic"
author: "Ben Best"
date: "August 9, 2016"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
suppressPackageStartupMessages({
  library(rerddap) # install.packages("rerddap")
  library(tidyverse)
  library(stringr)
  library(lubridate)
  library(DT)
  library(dygraphs)
  library(xts)
})
```

## Data

* [Marine Biodiversity Observation Network](http://mbon.ioos.us/)
* [Marine Biodiversity Observation Network: Florida Keys Reef Fish Visual Census](http://mbon.ioos.us/#module-metadata/53cb8d58-ae4d-43e1-aea3-640db9491db2/9436fa4e-73aa-4189-b551-b3b484a9d4e9)
    
    > This fish survey timeseries contains count and length measurements for 340 species observed between 1995 and 2012 along the southern coast of Florida.
    >
    > Filter options:
    > - Alpha: takes the average of event values within the selected area
    > - Gamma: groups all events within the selected area and treats them as a single sample
    > - Beta: Gamma/Alpha
    > 
    > Richness: Count of distinct species
    > - % Dominance (Berger-Parker): Numerical importance of the most abundant species
    > - Shannon-Wiener Diversity: This index quantifies the uncertainty associated with species prediction
    > - Pielou's Evenness: Species evenness quantifies how close in count each species is within a sampling event
    
    - [ERDDAP - Information about 1994 Florida Keys Reef Visual Census, v3.3, from Marine Biodiversity Observation Network](http://gcoos4.tamu.edu:8080/erddap/info/fk1994/index.html)

* [Marine Biodiversity Observation Network: SECOORA Fisheries and Wildlife Monitoring - Fisheries-Independent Monitoring, Assessment, and Prediction (FIMMAP)](http://mbon.ioos.us/#module-metadata/07a25110-e19e-4ad4-b29c-a6978b68a90d/200d32d2-c827-451d-846e-f3413090dbfa)
* [Marine Biodiversity Observation Network: SECOORA Fisheries and Wildlife Monitoring - Marine Resources Monitoring, Assessment, and Prediction (MARMAP)](http://mbon.ioos.us/#module-metadata/07a25110-e19e-4ad4-b29c-a6978b68a90d/7b846aaa-db05-408d-bd77-ae4f12281480)



* [RVC | Data Portal](http://www.sefsc.noaa.gov/rvc_analysis20/?acton=index)
* [jeremiaheb/rvc: New implementation of the rvcstats package, which produces summary statistics for the South Florida Reef Visual Census](https://github.com/jeremiaheb/rvc)


## rvc R package

- **REGION**: A code indicating the region. DRY TORT: Dry Tortugas, FLA KEYS: Florida Keys, and SEFCRI: Southeast Peninsular Florida.

```{r}
library(rvc) # devtools::install_github('jeremiaheb/rvc')

## Download desired Years/Regions data from server
fk = getRvcData(years = 1999:2014, regions='FLA KEYS') # c("FLA KEYS", "DRY TORT", "SEFCRI")

## Take a look at object structure
str(fk, 1) # 3 data.frames

## write taxa for manually adding groups
#write_csv(fk$taxonomic_data, 'data/rvc-fk_taxa.csv')

## Make a list of species
## You can use full scientific names, common names, or
## species codes (first 3 letters of genus, and first 4 of species)
## Only scientific names are case-sensitive
spcs = c("Epinephelus morio", "Black Grouper", "STE VARI")

## Calculate statistics for entire sampling domain
fk_dens = getDomainDensity(   fk, species = spcs)
fk_abun = getDomainAbundance( fk, species = spcs)
fk_occ  = getDomainOccurrence(fk, species = spcs)

## Stratum Level Estimates for Red Grouper in 2011 in
## The Florida Keys
sdens = getStratumDensity(keys11_14, species = "EPI MORI",
                         years = 2011, regions = "FLA KEYS")
```

```{r}
library(rvc) # devtools::install_github('jeremiaheb/rvc')

fk = getRvcData(years = 1999:2014, regions='FLA KEYS') # c("FLA KEYS", "DRY TORT", "SEFCRI")

spp = read_csv('data/rvc-fk_taxa.csv') %>%
  filter(group=='herbivores') # only flagged parrotfish

fk_dens = getDomainDensity(fk, species = spp$SCINAME)
fk_abun = getDomainAbundance(fk, species = spp$SCINAME)



# transform into xts format for dygraph

    
# density timeseries
fk_dens %>%
  filter(
    SPECIES_CD %in% c('SCA SPE.','SPA SPE.')) %>%
  mutate(
    #yr = as.Date(sprintf('%d-01-01', .$YEAR), format='%Y-%m-%d'),
    sp = recode(SPECIES_CD, 'SCA SPE.'='Scarus', 'SPA SPE.'='Sparisoma'),
    lwr = density - var,
    upr = density + var) %>%
  select(yr=YEAR, sp, density, lwr, upr) %>%
  gather(v, val, -yr, -sp) %>%
  unite(sp_v, sp, v) %>%
  spread(sp_v, val) %>%
  #as.xts(select(., -yr), order.by=.$yr) %>%
  dygraph(main='Herbivores (parrotfish) density in FL Keys') %>% 
  dySeries(c(   'Scarus_lwr',    'Scarus_density',    'Scarus_upr'), label =   'Scarus spp.') %>%
  dySeries(c('Sparisoma_lwr', 'Sparisoma_density', 'Sparisoma_upr'), label ='Sparisoma spp.') %>%
  dyRangeSelector()


# abundance timeseries
fk_abun %>%
  filter(
    SPECIES_CD %in% c('SCA SPE.','SPA SPE.')) %>%
  mutate(
    #yr = as.Date(sprintf('%d-01-01', .$YEAR), format='%Y-%m-%d'),
    sp = recode(SPECIES_CD, 'SCA SPE.'='Scarus', 'SPA SPE.'='Sparisoma'),
    lwr = max(abundance - 1.96 * var, 0),
    upr = abundance + 1.96 * var) %>%
  select(yr=YEAR, sp, abundance, lwr, upr) %>%
  gather(v, val, -yr, -sp) %>%
  unite(sp_v, sp, v) %>%
  spread(sp_v, val) %>%
  #as.xts(select(., -yr), order.by=.$yr) %>%
  dygraph(main='Herbivores (Parrotfish) Abundance in FL Keys') %>% 
  dySeries(c(   'Scarus_lwr',    'Scarus_abundance',    'Scarus_upr'), label =   'Scarus spp.') %>%
  dySeries(c('Sparisoma_lwr', 'Sparisoma_abundance', 'Sparisoma_upr'), label ='Sparisoma spp.') %>%
  dyRangeSelector()



# 

fk$sample_data %>%
  group_by(SPECIES_CD) %>%
  summarise(
    n = sum(NUM > 0)) %>%
  arrange(desc(n))
fk$sample_data %>%
  filter(SPECIES_CD == 'THA BIFA') %>% # ='Scarus'
  
  summary ise(SPECIES_NR)


Mean   : 0.03759

quantile(10)
mean()
quantile(90)
```



### LIVING RESOURCES

From Megan Hepner's "CR Questions and Datasets Available.docx":

**CR Question 7**: What is the status and trend of keystone and foundation species and how is it changing?

Potential indicators/data to answer the question	

-	Stony corals (CREMP data)
-	Long-spined sea urchin (RVC or CREMP data) - not in RVC
-	Herbivorous fish (RVC) 

- [NOAA CoRIS - Managing herbivores for their impacts on Caribbean coral reef ecosystems: A summary report for managers and practitioners](http://www.coris.noaa.gov/activities/herbivores_caribbean/)

    - Citation: Adam, T. C., Burkepile D. E., Ruttenberg B. I., Paddack M. J. (2015) [Managing herbivores for their impacts on Caribbean coral reef ecosystems: A summary report for managers and practitioners](http://data.nodc.noaa.gov/coris/library/NOAA/CRCP/project/182/Caribbean_herbivore_report_Final.pdf). PRBD-2015-1. US National Marine Fisheries Service
    
    Adam-etal_2015_Caribbean-Coral-Reef-Herbivores_Table1.csv

    - Reference: Adam TC, Burkepile DE, Ruttenberg BI, Paddack MJ (2015) [Herbivory and the resilience of Caribbean coral reefs: knowledge gaps and implications for management](http://www.coris.noaa.gov/exit.html?url=http%3A%2F%2Fwww.int-res.com%2Farticles%2Ffeature%2Fm520p001.pdf). Mar Ecol Prog Ser 520:1-20. doi: 10.3354/meps11170
    



Calculation/Metric 
-	Density 

**CR Question 8**: What is the status and trend of other focal species and how is it changing?

Potential indicators/data to answer the question	

-	Grunt-snapper complex (RVC)
    -	white grunt, blue stripped grunt, mutton snapper, gray snapper, yellowtail snapper, hogfish, red grouper, black grouper, Graysby
-	Queen conch (RVC)
-	Caribbean spiny lobster (RVC)
-	Sea turtles 

Calculation/Metric 
-	Density 

**CR Question 9**: What is the status of non-indigenous species and how is it changing?

Potential indicators/data to answer the question	

-	Lionfish (RVC, REEF, USGS data)
-	Orange cup coral (CREMP data)

Calculation/Metric 

-	Density 

**CR Question 10**: What is the status of biodiversity and how is it changing?

Potential indicators/data to answer the question	

-	Vertebrates (RVC/FIM)
-	Invertebrates (CREMP/RVC/FIM)

Calculation/Metric 

-	Species richness and evenness
-	Functional diversity 
-	Simpson index